### 2016-09-17 Thesis/DCG_sb-wi-thesis-template/MSpC/MSpC.tex
additional structuring
structuring
fixed bibliographic title capitalization by changing bst files
edits to intro
Changes to MSPC italics

The Morpheus MS search engine was recently designed for high-resolution, accurate-mass data obtained from Orbitrap-based instruments to provide faster matching of spectra to peptides \citep{wenger13}.
Unfortunately, no downstream automated tools are available to facilitate LFQ analysis, which can be quite challenging, if not impossible, to complete manually when accounting for shared peptides.
To overcome this bottleneck, we developed Morpheus Spectral Counter (MSpC) as the first LFQ computational tool that integrates directly with Morpheus to calculate NSAF, dNSAF, uNSAF, and corrected PSM \citep{fermin11} values in complex protein samples.
\begin{sidewaysfigure}[p]
	\centering
	\includegraphics[width=\columnwidth]{MSpC/figure1_supplemental.png}
	\mycaption{MSpC Graphic User Interface (GUI) and software flow chart}{
\textbf(A) Screenshot of the GUI.
The input requires the user to select a Morpheus search summary file containing experiments to be analyzed.  The user can optionally select a whitelist to filter output, and select an output directory.
Additional options can set peptide and protein FDR cutoffs, and method of quantification for output, including Normalized Spectral Abundance Factor (NSAF), distributed NSAF (dNSAF), and unique	NSAF (uNSAF).
A progress bar highlights completion of the analysis.
\textbf(B) Data analysis flow chart.
Experiments and groups of experiments to be analyzed are imported through the Morpheus summary file.
PSMs and protein groups are filtered at the specified FDR cutoff with a default of 1\%.}
	\label{fig:GUIflow}
\end{sidewaysfigure}
MSpC is fully automated, and only requires a Morpheus search summary file (summary.tsv) as input.
The user interface (see Figure \ref{fig:GUIflow} (A)) allows one to select the summary file and displays the raw MS/MS files that will be analyzed by MSpC.
Due to shared peptides being attributed to only one instance of a protein group in Morpheus's PSM file, PSMs are re-matched to all possible protein groups.
PSMs are then cataloged as shared or as unique (distinctly matching one protein group) to generate NSAF, dNSAF, and uNSAF outputs.
Finally, the output can be filtered for proteins of interest by specifying a comma delimited file containing unique identifiers and descriptions.
Some important features of MSpC are its ability to handle fractionation experiments as input, and the ability to whitelist proteins of interest in the output by specifying a csv file (see Tutorial \ref{sec:tutorial}).
Options exist to specify global PSM and protein group FDR rates (thus avoiding increased FDRs when one analyzes many experiments at once), to output NSAF, dNSAF, and uNSAF values, to require a minimum number of unique peptides to quantify a protein, and to specify an output directory.
A progress bar indicates completion of the analysis by MSpC.
To validate the accuracy of MSpC, we analyzed two MS/MS datasets available in PRIDE that were previously generated by high-energy collision-induced dissociation using Thermo Q-Exactive Orbitrap instruments.
\begin{figure}[p]
	\centering
	\includegraphics[width=\columnwidth]{MSpC/figure1.png}
	\mycaption{Confirmation of MSpC accuracy by analysis of MS/MS datasets generated with the Universal Proteome Standard 2 (UPS2)}{The array of UPS2 standards were spiked into Xenopus laevis egg \textbf{(Top)} and embryo \textbf{(Bottom)} extracts at a range of concentrations.  Following MS/MS analysis, dNSAF values for each protein were determined by Morpheus and MSpC.  \textbf{(A)} A log-log plot of dNSAF versus concentration for each UPS2 protein detected across each fmol range.  \textbf{(B)} A log-log plot of average dNSAF vs average concentration of each group of UPS2 proteins at each fmol range: (50, 500, 5000, and 50,000 fmol).}
	\label{fig:UPS2}
\end{figure}

Here, Xenopus egg (see top, Figure \ref{fig:UPS2}) and embryo (bottom, Figure \ref{fig:UPS2}) extracts were spiked at a 4:1 ratio with the Universal Proteome Standard 2 (UPS2), a mix of 48 purified proteins at defined molar ratios of 0.5, 5, 50, 500, 5000, and 50,000, with each ratio containing a different set of 8 of the 48 proteins.
As shown in Figure \ref{fig:UPS2} A, when the Morpheus/MSpC pipeline was used to calculate the average dNSAF value for each UPS2 protein, requiring only a single unique peptide to quantify, strong linear correlations (R$\textsuperscript{2}$ = 0.886 and 0.823) were obtained across a 1,000 fold change in abundance (50 fmol to 50,000 fmol).
In fact, the R$\textsuperscript{2}$ values were similar to those obtained by others with PSM-based LFQ methods \citep{cox14, tu14}.
This linear correlation was further strengthened when the dNSAF values were averaged for all UPS2 proteins within each of the concentration groups, with R$\textsuperscript{2}$ values of 0.994 and 0.992 for the egg and embryo datasets, respectively (see Figure \ref{fig:UPS2} B).
Notably, the slope of the concentration series was significantly less than unity, showing that NSAF measurements are not appropriate for absolute quantification, which was expected given that NSAF is a relative value.

@@

	\mycaption{Re- Analysis of MS/MS datasets generated with the Universal Proteome Standard 2 (UPS2)}{The array of UPS2 standards were spiked into Xenopus laevis egg \textbf{(Top)} and embryo \textbf{(Bottom)} extracts at a range of concentrations.  Following MS/MS analysis, dNSAF values for each protein were determined by Morpheus and MSpC with a change from Figure \ref{fig:UPS2} in that two unique peptides were required to quantify a protein. \textbf{(A)} A log-log plot of dNSAF versus concentration for each UPS2 protein detected across each fmol range.  \textbf{(B)} A log-log plot of average dNSAF vs average concentration of each group of UPS2 proteins at each fmol range: (50, 500, 5000, and 50,000 fmol).}

@@

Protein identifications were determined using the Morpheus search engine \citep{wenger13}.  Raw data was searched with the Thermo module of Morpheus revision 151 downloaded and compiled from source code available at \url{http://sourceforge.net/projects/morpheus-ms/} using Microsoft Visual Studio 2013 professional edition.  The following parameters were used to search all databases: unknown precursor charge states - +2, +3, +4; maximum number of MS/MS peaks = 400; assign charge states - enabled; de-isotope - disabled; generate target decoy database on the fly; protease trypsin (no proline rule); maximum missed cleavages = 2; initiator methionine behavior - variable; fixed modification of carbamidomethylation of cysteine; variable modification of oxidation of methionine; maximum variable modification isoforms per peptide = 1024; precursor mass tolerance = $\pm$ 2.1 Da monoisotopic (recommended parameters to account for neutral loss); precursor monoisotopic peak correction - disabled; product mass tolerance = $\pm$ 0.025 Da monoisotopic; consider modified forms as unique peptides - false; maximum threads = 8; minimize memory usage - false.   
MSpC quantification of Universal Proteome Standard 2 (UPS2) protein sequences exploited the MS/MS analysis of individual USP2 proteins (Sigma-Aldrich) mixed at various concentrations with egg or embryo extracts from Xenopus laevis available in the PRoteomics IDEntifications (PRIDE) repository \citep{vizcaino13} using identifier - PXD000902 and the available proteomics database - pita\_v1.71.protein.name.fa.  For our analysis, both the raw MS/MS data and resulting FASTA files for the USP2 and egg and embryo proteomes were obtained from PRIDE.  The database used for searching the MS/MS data of \textit{Arabidopsis} proteasomes spiked into total \textit{E. coli} peptides was generated by combining Uniprot K12 \textit{E. coli} reference proteome UP000000625 with a common contaminant database, and then mixing the merged dataset with FASTA sequences for all proteoforms of all known proteasome subunits and associated proteins \citep{book10} obtained from the TAIR10\_pep\_20101214 FASTA database available within The \textit{Arabidopsis} Information Resource (TAIR) version 10.  All FASTA files are available for download in the Supporting Information.  The datasets were analyzed by MSpC with a 1\% PSM false discovery rate (FDR) and a 1\% protein group FDR to determine NSAF, dNSAF, and uNSAF values for each protein group. Two separate analyses were performed in which one unique, or alternatively two unique peptides were required to quantify a protein group.  NSAF values were calculated according to formulas 1a, 2a, and 3a from Figure 1 of Zhang et al. \citep{zhang10}.

@@

\bibliographystyle{plantcell2}
\bibliography{MSpC}

>>>>

The Morpheus MS search engine was recently designed for high-resolution, accurate-mass data obtained from Orbitrap-based instruments to provide faster matching of spectra to peptides \citep{wenger13}.
Unfortunately, no downstream automated tools are available to facilitate LFQ analysis, which can be quite challenging, if not impossible, to complete manually when accounting for shared peptides.
To overcome this bottleneck, we developed Morpheus Spectral Counter (MSpC) as the first LFQ computational tool that integrates directly with Morpheus to calculate NSAF, dNSAF, uNSAF, and corrected PSM \citep{fermin11} values in complex protein samples.
\begin{sidewaysfigure}[p]
	\centering
	\includegraphics[width=\columnwidth]{MSpC/figure1_supplemental.png}
	\mycaption{MSpC Graphic User Interface (GUI) and software flow chart}{
\textbf(A) Screenshot of the GUI.
The input requires the user to select a Morpheus search summary file containing experiments to be analyzed.  The user can optionally select a whitelist to filter output, and select an output directory.
Additional options can set peptide and protein FDR cutoffs, and method of quantification for output, including Normalized Spectral Abundance Factor (NSAF), distributed NSAF (dNSAF), and unique	NSAF (uNSAF).
A progress bar highlights completion of the analysis.
\textbf(B) Data analysis flow chart.
Experiments and groups of experiments to be analyzed are imported through the Morpheus summary file.
PSMs and protein groups are filtered at the specified FDR cutoff with a default of 1\%.}
	\label{fig:GUIflow}
\end{sidewaysfigure}
MSpC is fully automated, and only requires a Morpheus search summary file (summary.tsv) as input.
The user interface (see Figure \ref{fig:GUIflow} (A)) allows one to select the summary file and displays the raw MS/MS files that will be analyzed by MSpC.
Due to shared peptides being attributed to only one instance of a protein group in Morpheus's PSM file, PSMs are re-matched to all possible protein groups.
PSMs are then cataloged as shared or as unique (distinctly matching one protein group) to generate NSAF, dNSAF, and uNSAF outputs.
Finally, the output can be filtered for proteins of interest by specifying a comma delimited file containing unique identifiers and descriptions.
Some important features of MSpC are its ability to handle fractionation experiments as input, and the ability to whitelist proteins of interest in the output by specifying a csv file (see Tutorial \ref{sec:tutorial}).
Options exist to specify global PSM and protein group FDR rates (thus avoiding increased FDRs when one analyzes many experiments at once), to output NSAF, dNSAF, and uNSAF values, to require a minimum number of unique peptides to quantify a protein, and to specify an output directory.
A progress bar indicates completion of the analysis by MSpC.
To validate the accuracy of MSpC, we analyzed two MS/MS datasets available in PRIDE that were previously generated by high-energy collision-induced dissociation using Thermo Q-Exactive Orbitrap instruments.
\begin{figure}[p]
	\centering
	\includegraphics[width=\columnwidth]{MSpC/figure1.png}
	\mycaption{Confirmation of MSpC accuracy by analysis of MS/MS datasets generated with the Universal Proteome Standard 2 (UPS2)}{The array of UPS2 standards were spiked into \textit{Xenopus laevis} egg \textbf{(Top)} and embryo \textbf{(Bottom)} extracts at a range of concentrations.  Following MS/MS analysis, dNSAF values for each protein were determined by Morpheus and MSpC.  \textbf{(A)} A log-log plot of dNSAF versus concentration for each UPS2 protein detected across each fmol range.  \textbf{(B)} A log-log plot of average dNSAF vs average concentration of each group of UPS2 proteins at each fmol range: (50, 500, 5000, and 50,000 fmol).}
	\label{fig:UPS2}
\end{figure}

Here, \textit{Xenopus} egg (see top, Figure \ref{fig:UPS2}) and embryo (bottom, Figure \ref{fig:UPS2}) extracts were spiked at a 4:1 ratio with the Universal Proteome Standard 2 (UPS2), a mix of 48 purified proteins at defined molar ratios of 0.5, 5, 50, 500, 5000, and 50,000, with each ratio containing a different set of 8 of the 48 proteins.
As shown in Figure \ref{fig:UPS2} A, when the Morpheus/MSpC pipeline was used to calculate the average dNSAF value for each UPS2 protein, requiring only a single unique peptide to quantify, strong linear correlations (R$\textsuperscript{2}$ = 0.886 and 0.823) were obtained across a 1,000 fold change in abundance (50 fmol to 50,000 fmol).
In fact, the R$\textsuperscript{2}$ values were similar to those obtained by others with PSM-based LFQ methods \citep{cox14, tu14}.
This linear correlation was further strengthened when the dNSAF values were averaged for all UPS2 proteins within each of the concentration groups, with R$\textsuperscript{2}$ values of 0.994 and 0.992 for the egg and embryo datasets, respectively (see Figure \ref{fig:UPS2} B).
Notably, the slope of the concentration series was significantly less than unity, showing that NSAF measurements are not appropriate for absolute quantification, which was expected given that NSAF is a relative value.

@@

	\mycaption{Re- Analysis of MS/MS datasets generated with the Universal Proteome Standard 2 (UPS2)}{The array of UPS2 standards were spiked into \textit{Xenopus laevis} egg \textbf{(Top)} and embryo \textbf{(Bottom)} extracts at a range of concentrations.  Following MS/MS analysis, dNSAF values for each protein were determined by Morpheus and MSpC with a change from Figure \ref{fig:UPS2} in that two unique peptides were required to quantify a protein. \textbf{(A)} A log-log plot of dNSAF versus concentration for each UPS2 protein detected across each fmol range.  \textbf{(B)} A log-log plot of average dNSAF vs average concentration of each group of UPS2 proteins at each fmol range: (50, 500, 5000, and 50,000 fmol).}

@@

Protein identifications were determined using the Morpheus search engine \citep{wenger13}.  Raw data was searched with the Thermo module of Morpheus revision 151 downloaded and compiled from source code available at \url{http://sourceforge.net/projects/morpheus-ms/} using Microsoft Visual Studio 2013 professional edition.  The following parameters were used to search all databases: unknown precursor charge states - +2, +3, +4; maximum number of MS/MS peaks = 400; assign charge states - enabled; de-isotope - disabled; generate target decoy database on the fly; protease trypsin (no proline rule); maximum missed cleavages = 2; initiator methionine behavior - variable; fixed modification of carbamidomethylation of cysteine; variable modification of oxidation of methionine; maximum variable modification isoforms per peptide = 1024; precursor mass tolerance = $\pm$ 2.1 Da monoisotopic (recommended parameters to account for neutral loss); precursor monoisotopic peak correction - disabled; product mass tolerance = $\pm$ 0.025 Da monoisotopic; consider modified forms as unique peptides - false; maximum threads = 8; minimize memory usage - false.   
MSpC quantification of Universal Proteome Standard 2 (UPS2) protein sequences exploited the MS/MS analysis of individual USP2 proteins (Sigma-Aldrich) mixed at various concentrations with egg or embryo extracts from \textit{Xenopus laevis} available in the PRoteomics IDEntifications (PRIDE) repository \citep{vizcaino13} using identifier - PXD000902 and the available proteomics database - pita\_v1.71.protein.name.fa.  For our analysis, both the raw MS/MS data and resulting FASTA files for the USP2 and egg and embryo proteomes were obtained from PRIDE.  The database used for searching the MS/MS data of \textit{Arabidopsis} proteasomes spiked into total \textit{E. coli} peptides was generated by combining Uniprot K12 \textit{E. coli} reference proteome UP000000625 with a common contaminant database, and then mixing the merged dataset with FASTA sequences for all proteoforms of all known proteasome subunits and associated proteins \citep{book10} obtained from the TAIR10\_pep\_20101214 FASTA database available within The \textit{Arabidopsis} Information Resource (TAIR) version 10.  All FASTA files are available for download in the Supporting Information.  The datasets were analyzed by MSpC with a 1\% PSM false discovery rate (FDR) and a 1\% protein group FDR to determine NSAF, dNSAF, and uNSAF values for each protein group. Two separate analyses were performed in which one unique, or alternatively two unique peptides were required to quantify a protein group.  NSAF values were calculated according to formulas 1a, 2a, and 3a from Figure 1 of Zhang et al. \citep{zhang10}.

@@

\begin{singlespace}
\bibliographystyle{plant_cell_final}
\renewcommand\bibname{Literature Cited}
\bibliography{MSpC}
\end{singlespace}
